---
title: "Untitled"
output: html_document
date: "2023-03-09"
---

```{r}
library(tsibble)
library(tsibbledata)
library(dplyr)
library(ggplot2)
library(feasts)
library(lubridate)
library(fpp3)
library(fpp2)
library(fpp)
library(padr)

```
### Handle data as tsibble

```{r}
head(ansett)

economy <- ansett %>%
  filter(Airports == "MEL-SYD" , Class == "Economy") %>%
  mutate(Passengers = Passengers / 1000)

head(economy)
```

```{r}
autoplot(economy , Passengers) +
  labs(title = "Ansett airlines economy class",
        subtitle = "Melbourne - Sydney",
        y = "Passengers ('000)")
```

```{r}
economy <- tsibble::fill_gaps(economy)

economy %>%
  gg_season(Passengers , labels = "right")
```

```{r}
economy %>%
  gg_subseries(Passengers)
```

### Australian tourist data

```{r}
holidays <- tourism %>%
  mutate(
    State = recode(State,
                   "Australian Capital Territory" = "ACT",
                   "New South Wales" = "NSW",
                   "Northern Territory" = "NT",
                   "Queensland" = "QLD"
    )
  ) %>%
  filter(Purpose == "Holiday") %>%
  group_by(State) %>%
  summarise(Trips = sum(Trips))

head(holidays)
```

#### Graph by States

```{r}
holidays %>%
  autoplot(Trips)
```

#### Seasonal plots

See the timings of seasonal peaks in each stats.

```{r}
holidays %>%
  gg_season(Trips)
```

#### Seasonal subseries plots

```{r}
holidays %>%
  gg_subseries(Trips)
```

Western Australian tourism has jumped marketedly in recent years, while Victorian tourism has increased in Q1 and Q4 but not in the middle of the year.

### Australian Beer production

#### line graph

```{r}
beer <- aus_production %>%
  select(Quarter , Beer)%>%
  filter(year(Quarter) >= 2000)

beer %>% autoplot(Beer)
```

#### Lagged plots

```{r}
beer %>%
  gg_lag(Beer , geom = "point" , lags = 1:9)
```

\*Relationships strongly positive at lags 4 and 8 -\> strong seasonality in the data\

\*Negative relationships seen for lags 2 and 6 -\>. peaks (in Q4) are plotted against troughs (in Q2)\

#### Calculate Autocorrelation

```{r}
beer %>%
  ACF(Beer, lag_max = 9)
```

### ACF plot (autocorrelation function plot)

```{r}
beer %>%
  ACF(Beer) %>%
  autoplot()
```

This plot shows autocorrelation at lags 1, 2,.. make up the autocorrelation function, ACF.\
The plot is also known as **correlogram**.

### Takeaway

**Trend** and **seasonality** in ACF plots:

-   When data have a trend, the auto correlations for small lags trend to be large and positive.

-   When data are seasonal, the auto correlations will be large at the seasonal lags (i.e. at multiples of the seasonal frequency)

-   When data are trended and seasonal, you see a combination of these effects.



```{r}
ts_all <- a10

ggplot(ts_all, aes(as.Date(ts_all), as.matrix(ts_all)))+ 
  geom_line(colour = "blue")+
  ylab("Sales") + 
  xlab("Date") 
```

```{r}
a10 <- fpp2::a10
a100 <- tsibble(Month = as.Date(a10),
               total_cost = c(as.matrix(a10)),
                index = Month,
               key = total_cost)

a100 %>%
  mutate(Month = yearmonth(Month)) %>%
  arrange(Month) -> a100

a100 %>%
  ACF(a100$total_cost , lag_max = 48) %>%
  autoplot()

a100 %>%
  autoplot(total_cost)

a100

```





